Detecting Hate Speech with GPT-3

### 2022-02-22 outputs/documents/paper.Rmd
Updates to results tables

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    toc: false
    fig_caption: true
    latex_engine: xelatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
  -  \usepackage{amsmath}
biblio-style: apalike
title: "Detecting Hate Speech with GPT-3"
thanks: "Code and data are available at: https://github.com/kelichiu/GPT3-hate-speech-detection. We gratefully acknowledge the support of Gillian Hadfield, the Schwartz Reisman Institute for Technology and Society, and OpenAI for providing access to GPT-3 under the academic access program. We thank Amy Farrow, Haoluan Chen, John Giorgi, Mauricio Vargas Sepúlveda, Monica Alexander, Noam Kolt, and Tom Davidson for helpful discussions and suggestions. Please note that we have added asterisks to racial slurs and other offensive content in this paper, however the inputs and outputs did not have these.  Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto and Schwartz Reisman Institute
abstract: "Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist? We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 can identify sexist or racist text with an accuracy between 48 per cent and 69 per cent. With few-shot learning and an instruction included in the prompt, the model's accuracy can be as high as 78 per cent. We conclude that large language models have a role to play in hate speech detection, and that with further development language models could be used to counter hate speech and even self-police."
keywords: "GPT-3; natural language processing; quantitative analysis; hate speech."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
endnote: no
graphics: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), echo = FALSE, warning = FALSE)
# library(dplyr)
library(tidyverse)
# library(tidyr)
# library(readr)
```

@@

```{r zeroshot}
zero_shot_result <- read_csv(here::here("outputs/data/zero_shot_results.csv"), show_col_types = FALSE)

zero_shot_result <- 
  zero_shot_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  mutate(result = case_when(answer_yn==label_yn ~ "Match",
                            answer_yn!=label_yn ~ "Mismatch"
                            )) %>%
  group_by(result, category) %>%
  summarise(result_cnt = n(), .groups="drop") %>% 
  mutate(category = case_when(category == "racist" ~ "Racist",
                              category == "sexist" ~ "Sexist"
                              )) %>%
  mutate(Percent = result_cnt / sum(result_cnt) * 100,
         Percent = round(Percent)) %>%
  rename(Result = result,
         Category = category,
         Count = result_cnt)

zero_shot_result %>% 
  knitr::kable(caption = "Classification of statements with zero-shot learning",
               booktabs = TRUE)
```

@@

```{r oneshot}
one_shot_result <- read.csv(here::here("outputs/data/one_shot_results.csv"))

one_shot_result <- 
  one_shot_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  mutate(result = case_when(answer_yn==label_yn ~ "Match",
                            answer_yn!=label_yn ~ "Mismatch"
                            )) %>%
  group_by(result, category) %>%
  summarise(result_cnt = n(), .groups = "drop") %>% 
  mutate(category = case_when(category == "racist" ~ "Racist",
                              category == "sexist" ~ "Sexist"
                              )) %>%
  mutate(Percent = result_cnt / sum(result_cnt) * 100,
         Percent = round(Percent)) %>%
  rename(Result = result,
         Category = category,
         Count = result_cnt)

one_shot_result %>% 
  knitr::kable(caption = "Classification of statements with one-shot learning", 
             booktabs = TRUE)
```

@@

```{r fewshotsingle}
few_shot_single_result <- read.csv(here::here("outputs/data/few_shot_single_results.csv"))

few_shot_single_result <- 
  few_shot_single_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer) %>%
  mutate(label_yn = ifelse(grepl("not", label), "No", "Yes")) %>%
  mutate(answer_yn = ifelse(grepl("not", answer), "No", "Yes")) %>%
  mutate(result = case_when(answer_yn==label_yn ~ "Match",
                            answer_yn!=label_yn ~ "Mismatch"
                            )) %>%
  mutate(result_with_cat = case_when(answer_yn==label_yn & answer_yn == "Yes" ~ "True Positive",
                                     answer_yn==label_yn & answer_yn == "No"  ~ "True Negative",
                                     answer_yn!=label_yn & answer_yn == "Yes" ~ "False Positive",
                                     answer_yn!=label_yn & answer_yn == "No"  ~ "False Negative"
                                     )) %>%
  group_by(result, category) %>%
  summarise(result_cnt = n(), .groups = "drop") %>% 
  mutate(category = case_when(category == "racist" ~ "Racist",
                              category == "sexist" ~ "Sexist"
                              )) %>%
  mutate(Percent = result_cnt / sum(result_cnt) * 100,
         Percent = round(Percent)) %>%
  rename(Result = result,
         Category = category,
         Count = result_cnt)

few_shot_single_result %>% 
  knitr::kable(
    caption = "Classification of statements with single-category few-shot learning", 
    booktabs = TRUE)
```

>>>>

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    toc: false
    fig_caption: true
    latex_engine: xelatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
  -  \usepackage{amsmath}
biblio-style: apalike
title: "Detecting Hate Speech with GPT-3"
thanks: "Code and data are available at: https://github.com/kelichiu/GPT3-hate-speech-detection. We gratefully acknowledge the support of Gillian Hadfield, the Schwartz Reisman Institute for Technology and Society, and OpenAI for providing access to GPT-3 under the academic access program. We thank Amy Farrow, Haoluan Chen, John Giorgi, Mauricio Vargas Sepúlveda, Monica Alexander, Noam Kolt, and Tom Davidson for helpful discussions and suggestions. Please note that we have added asterisks to racial slurs and other offensive content in this paper, however the inputs and outputs did not have these.  Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto and Schwartz Reisman Institute
abstract: "Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist? We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 can identify sexist or racist text with an accuracy between 48 per cent and 69 per cent. With few-shot learning and an instruction included in the prompt, the model's accuracy can be as high as 78 per cent. We conclude that large language models have a role to play in hate speech detection, and that with further development language models could be used to counter hate speech and even self-police."
keywords: "GPT-3; natural language processing; quantitative analysis; hate speech."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
endnote: no
graphics: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), echo = FALSE, warning = FALSE)
# library(dplyr)
library(tidyverse)
library(kableExtra)
# library(tidyr)
# library(readr)
```

@@

<!-- ```{r zeroshot} -->
<!-- zero_shot_result <- read_csv(here::here("outputs/data/zero_shot_results.csv"), show_col_types = FALSE) -->

<!-- zero_shot_result_match <-  -->
<!--   zero_shot_result %>% -->
<!--   filter(temperature == 0) %>% -->
<!--   select(category, label, answer, temperature) %>% -->
<!--   mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>% -->
<!--   mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>% -->
<!--   mutate(result = case_when(answer_yn==label_yn ~ "Match", -->
<!--                             answer_yn!=label_yn ~ "Mismatch" -->
<!--                             )) %>% -->
<!--   group_by(result, category) %>% -->
<!--   summarise(result_cnt = n(), .groups="drop") %>%  -->
<!--   mutate(category = case_when(category == "racist" ~ "Racist", -->
<!--                               category == "sexist" ~ "Sexist" -->
<!--                               )) %>% -->
<!--   mutate(Percent = result_cnt / sum(result_cnt) * 100, -->
<!--          Percent = round(Percent)) %>% -->
<!--   rename(Result = result, -->
<!--          Category = category, -->
<!--          Count = result_cnt) -->

<!-- zero_shot_result_match %>%  -->
<!--   knitr::kable(caption = "Classification of statements with zero-shot learning", -->
<!--                booktabs = TRUE) -->
<!-- ``` -->

```{r zeroshot-racism}
zero_shot_result <- read_csv(here::here("outputs/data/zero_shot_results.csv"), show_col_types = FALSE)

zero_shot_result_racism <- 
  zero_shot_result %>%
  filter(temperature == 0 & category == "racist") %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  count(label_yn, answer_yn) %>%
  pivot_wider(names_from = answer_yn, values_from = n, values_fill = 0) %>%
  mutate(label_yn = case_when(label_yn == "N" ~ "Not racist",
                              label_yn == "Y" ~ "Racist")) %>%
  rename("Actual classification" = label_yn, "Not racist" = N, "Racist" = Y)

zero_shot_result_racism <- zero_shot_result_racism[, c("Actual classification", "Not racist", "Racist")]

zero_shot_result_racism %>%
  knitr::kable(caption = "Classification of racist statements with zero-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2))

```

```{r zeroshot-sexism}
zero_shot_result_sexism <- 
  zero_shot_result %>%
  filter(temperature == 0 & category == "sexist") %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  count(label_yn, answer_yn) %>%
  pivot_wider(names_from = answer_yn, values_from = n, values_fill = 0) %>%
  mutate(label_yn = case_when(label_yn == "N" ~ "Not sexist",
                              label_yn == "Y" ~ "Sexist")) %>%
  rename("Actual classification" = label_yn, "Not sexist" = N, "Sexist" = Y)

zero_shot_result_sexism <- zero_shot_result_sexism[, c("Actual classification", "Not sexist", "Sexist")]

zero_shot_result_sexism %>%
  knitr::kable(caption = "Classification of sexist statements with zero-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2))
```

```{r zeroshot-hate}
zero_shot_result_hate <- 
  zero_shot_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  count(label_yn, answer_yn) %>%
  pivot_wider(names_from = answer_yn, values_from = n, values_fill = 0) %>%
  mutate(label_yn = case_when(label_yn == "N" ~ "Not hate speech",
                              label_yn == "Y" ~ "Hate speech")) %>%
  rename("Actual classification" = label_yn, "Not hate speech" = N, "Hate speech" = Y)

zero_shot_result_hate <- zero_shot_result_hate[, c("Actual classification", "Not hate speech", "Hate speech")]

zero_shot_result_hate %>%
  knitr::kable(caption = "Classification of hate speech with zero-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2))
```

```{r zeroshot-summary}
tp_r <- zero_shot_result_racism[2, 3] %>% as.numeric()
tn_r <- zero_shot_result_racism[1, 2] %>% as.numeric()
fp_r <- zero_shot_result_racism[1, 3] %>% as.numeric()
fn_r <- zero_shot_result_racism[2, 2] %>% as.numeric()

tp_s <- zero_shot_result_sexism[2, 3] %>% as.numeric()
tn_s <- zero_shot_result_sexism[1, 2] %>% as.numeric()
fp_s <- zero_shot_result_sexism[1, 3] %>% as.numeric()
fn_s <- zero_shot_result_sexism[2, 2] %>% as.numeric()

tp_h <- zero_shot_result_hate[2, 3] %>% as.numeric()
tn_h <- zero_shot_result_hate[1, 2] %>% as.numeric()
fp_h <- zero_shot_result_hate[1, 3] %>% as.numeric()
fn_h <- zero_shot_result_hate[2, 2] %>% as.numeric()

zero_racism_accuracy <- ((tp_r + tn_r)/(tp_r + tn_r + fp_r + fn_r)) %>% round(4)*100
zero_racism_precision <- (tp_r/(tp_r + fp_r)) %>% round(4)*100
zero_racism_recall <- (tp_r/(tp_r + fn_r)) %>% round(4)*100
zero_racism_f1 <- ((2*zero_racism_precision*zero_racism_recall)/(zero_racism_precision + zero_racism_recall)) %>% round(2)
zero_racism_all <- c(zero_racism_accuracy, zero_racism_precision, zero_racism_recall, zero_racism_f1)

zero_sexism_accuracy <- ((tp_s + tn_s)/(tp_s + tn_s + fp_s + fn_s)) %>% round(4)*100
zero_sexism_precision <- (tp_s/(tp_s + fp_s)) %>% round(4)*100
zero_sexism_recall <- (tp_s/(tp_s + fn_s)) %>% round(4)*100
zero_sexism_f1 <- ((2*zero_sexism_precision*zero_sexism_recall)/(zero_sexism_precision + zero_sexism_recall)) %>% round(2)
zero_sexism_all <- c(zero_sexism_accuracy, zero_sexism_precision, zero_sexism_recall, zero_sexism_f1)

zero_hate_accuracy <- ((tp_h + tn_h)/(tp_h + tn_h + fp_h + fn_h)) %>% round(4)*100
zero_hate_precision <- (tp_h/(tp_h + fp_h)) %>% round(4)*100
zero_hate_recall <- (tp_h/(tp_h + fn_h)) %>% round(4)*100
zero_hate_f1 <- ((2*zero_hate_precision*zero_sexism_recall)/(zero_hate_precision + zero_hate_recall)) %>% round(2)
zero_hate_all <- c(zero_hate_accuracy, zero_hate_precision, zero_hate_recall, zero_hate_f1)

zero_shot_summary <- tibble(Category = c(rep("Racism", 4), rep("Sexism", 4), rep("Overall", 4)), 
                                   Metric = rep(c("Accuracy", "Precision", "Recall", "F1"), 3), 
                                   Percent = c(zero_racism_all, zero_sexism_all, zero_hate_all))

zero_shot_summary %>%
  select(!Category) %>%
  knitr::kable(caption="Performance of zero-shot learning in hate speech classification",
               col.names=c("", "Percent"),
               booktabs=TRUE) %>%
  kableExtra::pack_rows(index = c("Racism"=4, "Sexism"=4, "Overall"=4))
```

@@

<!-- ```{r oneshot} -->
<!-- one_shot_result <- read.csv(here::here("outputs/data/one_shot_results.csv")) -->

<!-- one_shot_result <-  -->
<!--   one_shot_result %>% -->
<!--   filter(temperature == 0) %>% -->
<!--   select(category, label, answer) %>% -->
<!--   mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>% -->
<!--   mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>% -->
<!--   mutate(result = case_when(answer_yn==label_yn ~ "Match", -->
<!--                             answer_yn!=label_yn ~ "Mismatch" -->
<!--                             )) %>% -->
<!--   group_by(result, category) %>% -->
<!--   summarise(result_cnt = n(), .groups = "drop") %>%  -->
<!--   mutate(category = case_when(category == "racist" ~ "Racist", -->
<!--                               category == "sexist" ~ "Sexist" -->
<!--                               )) %>% -->
<!--   mutate(Percent = result_cnt / sum(result_cnt) * 100, -->
<!--          Percent = round(Percent)) %>% -->
<!--   rename(Result = result, -->
<!--          Category = category, -->
<!--          Count = result_cnt) -->

<!-- one_shot_result %>%  -->
<!--   knitr::kable(caption = "Classification of statements with one-shot learning",  -->
<!--              booktabs = TRUE) -->
<!-- ``` -->

```{r oneshot-racism}
one_shot_result <- read.csv(here::here("outputs/data/one_shot_results.csv"))

one_shot_result_racism <- 
  one_shot_result %>%
  filter(temperature == 0 & category == "racist") %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  count(label_yn, answer_yn) %>%
  pivot_wider(names_from = answer_yn, values_from = n, values_fill = 0) %>%
  mutate(label_yn = case_when(label_yn == "N" ~ "Not racist",
                              label_yn == "Y" ~ "Racist")) %>%
  rename("Actual classification" = label_yn, "Not racist" = N, "Racist" = Y)

one_shot_result_racism <- one_shot_result_racism[, c("Actual classification", "Not racist", "Racist")]

one_shot_result_racism %>%
  knitr::kable(caption = "Classification of racist statements with one-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2))

```

```{r oneshot-sexism}
one_shot_result_sexism <- 
  one_shot_result %>%
  filter(temperature == 0 & category == "sexist") %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  count(label_yn, answer_yn) %>%
  pivot_wider(names_from = answer_yn, values_from = n, values_fill = 0) %>%
  mutate(label_yn = case_when(label_yn == "N" ~ "Not sexist",
                              label_yn == "Y" ~ "Sexist")) %>%
  rename("Actual classification" = label_yn, "Not sexist" = N, "Sexist" = Y)

one_shot_result_sexism <- one_shot_result_sexism[, c("Actual classification", "Not sexist", "Sexist")]

one_shot_result_sexism %>%
  knitr::kable(caption = "Classification of sexist statements with one-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2))
```

```{r oneshot-hate}
one_shot_result_hate <- 
  one_shot_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  count(label_yn, answer_yn) %>%
  pivot_wider(names_from = answer_yn, values_from = n, values_fill = 0) %>%
  mutate(label_yn = case_when(label_yn == "N" ~ "Not hate speech",
                              label_yn == "Y" ~ "Hate speech")) %>%
  rename("Actual classification" = label_yn, "Not hate speech" = N, "Hate speech" = Y)

one_shot_result_hate <- one_shot_result_hate[, c("Actual classification", "Not hate speech", "Hate speech")]

one_shot_result_hate %>%
  knitr::kable(caption = "Classification of hate speech with one-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2))
```

```{r oneshot-summary}
tp_r <- one_shot_result_racism[2, 3] %>% as.numeric()
tn_r <- one_shot_result_racism[1, 2] %>% as.numeric()
fp_r <- one_shot_result_racism[1, 3] %>% as.numeric()
fn_r <- one_shot_result_racism[2, 2] %>% as.numeric()

tp_s <- one_shot_result_sexism[2, 3] %>% as.numeric()
tn_s <- one_shot_result_sexism[1, 2] %>% as.numeric()
fp_s <- one_shot_result_sexism[1, 3] %>% as.numeric()
fn_s <- one_shot_result_sexism[2, 2] %>% as.numeric()

tp_h <- one_shot_result_hate[2, 3] %>% as.numeric()
tn_h <- one_shot_result_hate[1, 2] %>% as.numeric()
fp_h <- one_shot_result_hate[1, 3] %>% as.numeric()
fn_h <- one_shot_result_hate[2, 2] %>% as.numeric()

one_racism_accuracy <- ((tp_r + tn_r)/(tp_r + tn_r + fp_r + fn_r)) %>% round(4)*100
one_racism_precision <- (tp_r/(tp_r + fp_r)) %>% round(4)*100
one_racism_recall <- (tp_r/(tp_r + fn_r)) %>% round(4)*100
one_racism_f1 <- ((2*one_racism_precision*one_racism_recall)/(one_racism_precision + one_racism_recall)) %>% round(2)
one_racism_all <- c(one_racism_accuracy, one_racism_precision, one_racism_recall, one_racism_f1)

one_sexism_accuracy <- ((tp_s + tn_s)/(tp_s + tn_s + fp_s + fn_s)) %>% round(4)*100
one_sexism_precision <- (tp_s/(tp_s + fp_s)) %>% round(4)*100
one_sexism_recall <- (tp_s/(tp_s + fn_s)) %>% round(4)*100
one_sexism_f1 <- ((2*one_sexism_precision*one_sexism_recall)/(one_sexism_precision + one_sexism_recall)) %>% round(2)
one_sexism_all <- c(one_sexism_accuracy, one_sexism_precision, one_sexism_recall, one_sexism_f1)

one_hate_accuracy <- ((tp_h + tn_h)/(tp_h + tn_h + fp_h + fn_h)) %>% round(4)*100
one_hate_precision <- (tp_h/(tp_h + fp_h)) %>% round(4)*100
one_hate_recall <- (tp_h/(tp_h + fn_h)) %>% round(4)*100
one_hate_f1 <- ((2*one_hate_precision*one_sexism_recall)/(one_hate_precision + one_hate_recall)) %>% round(2)
one_hate_all <- c(one_hate_accuracy, one_hate_precision, one_hate_recall, one_hate_f1)

one_shot_summary <- tibble(Category = c(rep("Racism", 4), rep("Sexism", 4), rep("Overall", 4)), 
                                   Metric = rep(c("Accuracy", "Precision", "Recall", "F1"), 3), 
                                   Percent = c(one_racism_all, one_sexism_all, one_hate_all))

one_shot_summary %>%
  select(!Category) %>%
  knitr::kable(caption="Performance of one-shot learning in hate speech classification",
               col.names=c("", "Percent"),
               booktabs=TRUE) %>%
  kableExtra::pack_rows(index = c("Racism"=4, "Sexism"=4, "Overall"=4))
```

@@

<!-- ```{r fewshotsingle} -->
<!-- few_shot_single_result <- read.csv(here::here("outputs/data/few_shot_single_results.csv")) -->

<!-- few_shot_single_result <-  -->
<!--   few_shot_single_result %>% -->
<!--   filter(temperature == 0) %>% -->
<!--   select(category, label, answer) %>% -->
<!--   mutate(label_yn = ifelse(grepl("not", label), "No", "Yes")) %>% -->
<!--   mutate(answer_yn = ifelse(grepl("not", answer), "No", "Yes")) %>% -->
<!--   mutate(result = case_when(answer_yn==label_yn ~ "Match", -->
<!--                             answer_yn!=label_yn ~ "Mismatch" -->
<!--                             )) %>% -->
<!--   mutate(result_with_cat = case_when(answer_yn==label_yn & answer_yn == "Yes" ~ "True Positive", -->
<!--                                      answer_yn==label_yn & answer_yn == "No"  ~ "True Negative", -->
<!--                                      answer_yn!=label_yn & answer_yn == "Yes" ~ "False Positive", -->
<!--                                      answer_yn!=label_yn & answer_yn == "No"  ~ "False Negative" -->
<!--                                      )) %>% -->
<!--   group_by(result, category) %>% -->
<!--   summarise(result_cnt = n(), .groups = "drop") %>%  -->
<!--   mutate(category = case_when(category == "racist" ~ "Racist", -->
<!--                               category == "sexist" ~ "Sexist" -->
<!--                               )) %>% -->
<!--   mutate(Percent = result_cnt / sum(result_cnt) * 100, -->
<!--          Percent = round(Percent)) %>% -->
<!--   rename(Result = result, -->
<!--          Category = category, -->
<!--          Count = result_cnt) -->

<!-- few_shot_single_result %>%  -->
<!--   knitr::kable( -->
<!--     caption = "Classification of statements with single-category few-shot learning",  -->
<!--     booktabs = TRUE) -->
<!-- ``` -->

```{r fewshotsingle-racism}
few_shot_single_result <- read.csv(here::here("outputs/data/few_shot_single_results.csv"))

few_shot_result_racism <- 
  few_shot_single_result %>%
  filter(temperature == 0 & category == "racist") %>%
  select(category, label, answer) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "not", negate = FALSE), "N", "Y")) %>%
  count(label_yn, answer_yn) %>%
  pivot_wider(names_from = answer_yn, values_from = n, values_fill = 0) %>%
  mutate(label_yn = case_when(label_yn == "N" ~ "Not racist",
                              label_yn == "Y" ~ "Racist")) %>%
  rename("Actual classification" = label_yn, "Not racist" = N, "Racist" = Y)

few_shot_result_racism <- few_shot_result_racism[, c("Actual classification", "Not racist", "Racist")]

few_shot_result_racism %>%
  knitr::kable(caption = "Classification of racist statements with single-category few-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2))
```

```{r fewshotsingle-sexism}
few_shot_result_sexism <- 
  few_shot_single_result %>%
  filter(temperature == 0 & category == "sexist") %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "not", negate = FALSE), "N", "Y")) %>%
  count(label_yn, answer_yn) %>%
  pivot_wider(names_from = answer_yn, values_from = n, values_fill = 0) %>%
  mutate(label_yn = case_when(label_yn == "N" ~ "Not sexist",
                              label_yn == "Y" ~ "Sexist")) %>%
  rename("Actual classification" = label_yn, "Not sexist" = N, "Sexist" = Y)

few_shot_result_sexism <- few_shot_result_sexism[, c("Actual classification", "Not sexist", "Sexist")]

few_shot_result_sexism %>%
  knitr::kable(caption = "Classification of sexist statements with single-category few-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2))
```

```{r fewshotsingle-hate}
few_shot_result_hate <- 
  few_shot_single_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "not", negate = FALSE), "N", "Y")) %>%
  count(label_yn, answer_yn) %>%
  pivot_wider(names_from = answer_yn, values_from = n, values_fill = 0) %>%
  mutate(label_yn = case_when(label_yn == "N" ~ "Not hate speech",
                              label_yn == "Y" ~ "Hate speech")) %>%
  rename("Actual classification" = label_yn, "Not hate speech" = N, "Hate speech" = Y)

few_shot_result_hate <- few_shot_result_hate[, c("Actual classification", "Not hate speech", "Hate speech")]

few_shot_result_hate %>%
  knitr::kable(caption = "Classification of hate speech with single-category few-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2))
```

```{r fewshotsingle-summary}
tp_r <- few_shot_result_racism[2, 3] %>% as.numeric()
tn_r <- few_shot_result_racism[1, 2] %>% as.numeric()
fp_r <- few_shot_result_racism[1, 3] %>% as.numeric()
fn_r <- few_shot_result_racism[2, 2] %>% as.numeric()

tp_s <- few_shot_result_sexism[2, 3] %>% as.numeric()
tn_s <- few_shot_result_sexism[1, 2] %>% as.numeric()
fp_s <- few_shot_result_sexism[1, 3] %>% as.numeric()
fn_s <- few_shot_result_sexism[2, 2] %>% as.numeric()

tp_h <- few_shot_result_hate[2, 3] %>% as.numeric()
tn_h <- few_shot_result_hate[1, 2] %>% as.numeric()
fp_h <- few_shot_result_hate[1, 3] %>% as.numeric()
fn_h <- few_shot_result_hate[2, 2] %>% as.numeric()

few_racism_accuracy <- ((tp_r + tn_r)/(tp_r + tn_r + fp_r + fn_r)) %>% round(4)*100
few_racism_precision <- (tp_r/(tp_r + fp_r)) %>% round(4)*100
few_racism_recall <- (tp_r/(tp_r + fn_r)) %>% round(4)*100
few_racism_f1 <- ((2*few_racism_precision*few_racism_recall)/(few_racism_precision + few_racism_recall)) %>% round(2)
few_racism_all <- c(few_racism_accuracy, few_racism_precision, few_racism_recall, few_racism_f1)

few_sexism_accuracy <- ((tp_s + tn_s)/(tp_s + tn_s + fp_s + fn_s)) %>% round(4)*100
few_sexism_precision <- (tp_s/(tp_s + fp_s)) %>% round(4)*100
few_sexism_recall <- (tp_s/(tp_s + fn_s)) %>% round(4)*100
few_sexism_f1 <- ((2*few_sexism_precision*few_sexism_recall)/(few_sexism_precision + few_sexism_recall)) %>% round(2)
few_sexism_all <- c(few_sexism_accuracy, few_sexism_precision, few_sexism_recall, few_sexism_f1)

few_hate_accuracy <- ((tp_h + tn_h)/(tp_h + tn_h + fp_h + fn_h)) %>% round(4)*100
few_hate_precision <- (tp_h/(tp_h + fp_h)) %>% round(4)*100
few_hate_recall <- (tp_h/(tp_h + fn_h)) %>% round(4)*100
few_hate_f1 <- ((2*few_hate_precision*few_sexism_recall)/(few_hate_precision + few_hate_recall)) %>% round(2)
few_hate_all <- c(few_hate_accuracy, few_hate_precision, few_hate_recall, few_hate_f1)

few_shot_summary <- tibble(Category = c(rep("Racism", 4), rep("Sexism", 4), rep("Overall", 4)), 
                                   Metric = rep(c("Accuracy", "Precision", "Recall", "F1"), 3), 
                                   Percent = c(few_racism_all, few_sexism_all, few_hate_all))

few_shot_summary %>%
  select(!Category) %>%
  knitr::kable(caption="Performance of single-category few-shot learning in hate speech classification",
               col.names=c("", "Percent"),
               booktabs=TRUE) %>%
  kableExtra::pack_rows(index = c("Racism"=4, "Sexism"=4, "Overall"=4))
```
